Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-08-06 volunteer-clouds-ieee-ci-si.tex
Response to comments
Trabajaré en explicar la figura y la implementación con web workers.
Using both figures from @mariosky and @fergunet
Helping to describe the table

Some time ago, our group faced the problem of diminishing funds for
buying new     % [Pedro] This sentence is so so long. I would split it
               % in two sentences. [JJ] Done.
hardware. This was aggravated by the increasing maintenance costs and extended downtime resulting from the continuous failures of existing clusters.
%[Mario] Otra opción: Some time ago, our group faced the problem of diminishing funds 
%for buying new hardware. This was aggravated by the increasing maintance costs 
%resulting from the continuous failures of existing clusters.
% [JJ] Done 
Considering this, we leveraged our experience in the design of web
applications, including JavaScript, since their inception and other
volunteer and 
unconventional distributed evolutionary computing systems, and we 
decided to design and release a new free framework that would allow anyone to
create a volunteer distributed EC experiment using cloud resources as
servers and browsers as clients.

@@

%Algo novedoso es el uso de Javascript no solo en el cliente si no también en el Servidor.
%Como el tema es el Software creo que esto se tiene que resaltar. En los sistemas anteriores 
%se tenía lo de Java en el cliente, pero en el server estaba lo grueso del trabajo:
% Administrar el server - Ahora más fácil con cloud
% Programar el server side - Ahora se puede usar JavaScript también para eso.

@@

Figure \ref{fig:system} describes the general system architecture and
algorithm behavior. Different web technologies, such as JQuery or chart.js have
been used to build the elements of the system. 
%FERGU: Mario, explica en el texto de abajo qué es winston en la figura también. 
%Vamos, que todo lo que salga en la figura tiene que ir en el texto en alguna parte :)
%
\begin{figure}[!t]
\centering
\includegraphics[width=4in]{img/Algorithm.pdf}
\caption{Description of the proposed system. Clients execute a JS EA
  in the browser, which, every 100 generations, sends the best
  individual and receives a random one from the server. }  
\label{fig:system}
\end{figure}

@@

\begin{figure}[!t]
\centering
\includegraphics[width=12cm]{img/baseline-times.png}
\caption{Comparison of time to solution for the baseline system, using a {\tt
    node.js} client, for different population sizes. They only compare on
runs in which solution has been found.}
\label{fig:baseline}
\end{figure}
First we establish the baseline performance by running the
evolutionary algorithm in a desktop client written using NodEO, 
%FERGU: es NodEO o NodIO? Quito el \sf, por cierto, que queda feo
%Por qué usamos NodEO aquí y no NodIO, vamos a compararlos? 
%Qué diferencias hay entre los dos? (ponerlo en el SOA)
the node.js evolutionary algorithm library. This experiment tries to 
%FERGU: quitar el "the node.js evolutionary algorithm library", si ya lo hemos dicho
%o explicarlo mejor: "the basic js library we have used as base to built NodIO", por ejemplo
find the solution to the 40-trap function with parameters $l=4, a=1,
b=2, z=3$, a population size equal 512, which was found in previous
experiments to be the correct one for finding the solution 95\% of the
%Fergu: cita a esos experimentos?
time. The algorithm was run until the solution (a string with all
ones) was found or five millions evaluations had been performed. It
took around a minute, on average, that is $68.9694$ seconds, to
perform the fifty runs. In this experiment only 33, that is, 66\% were
successful. The experiments were repeated for population $p=1024$,
with success rate upgraded to 100\% and average duration $3.46$
seconds. Results for these two experiments 
are shown in Figure \ref{fig:baseline}.

@@

Initial experiments were set up using the OpenShift
Platform-as-a-Service, which provides a free tier within which this
experiment could be performed. That way, and as required, the hardware
and cloud cost for this experiment were zero. Experiments were
announced through a post in Twitter and other social networks, and
results were published here \cite{DBLP:conf/gecco/GuervosG15}. For the
purpose of this paper, we repeated the announcement several times
through the months of April and then by the beginning of August. All
in all, we have the set of runs with the characteristics shown in
Table \ref{tab:runs}.
%
\begin{table*}
\caption{Characteristics of the volunteer computing initial runs. \label{tab:runs}}
\begin{center}
\begin{tabular}{l|rr}
\hline
Date & \# experiments & Different IPs \\
\hline 
April 4th 4/4 & 57 & 191 \\
April 24th 4/24 &  231 & 559 \\
July 31th 7/31 & 97 & 179 \\
\hline 
\end{tabular}
\end{center}
\end{table*}
%
\begin{table*}
\caption{Summary of time per run, number of IPS and number of {\tt
    PUT}s per IP in the initial runs. \label{tab:summary}}
\begin{center}
\begin{tabular}{l|ccccc}
\hline
Date & Median \#IPs & Median time (s) & Median \#{\tt PUT}s & $<$ 69s & $<$ 3.46s \\
\hline 
4/4 & 5 & 2040 & 18 & 14.29\% & 5.36\% \\
4/24 &  5 & 732 & 11 & 47.39\% & 3.91\% \\
7/31 & 5 & 260 & 23 & 27.08\% & 1.04\%  \\
\hline 
\end{tabular}
\end{center}
\end{table*}
%
A summary of the characteristics of each run is also shown in Table
\ref{tab:summary}.

>>>>

Some time ago, our group faced the problem of diminishing funds for
buying new     % [Pedro] This sentence is so so long. I would split it
               % in two sentences. [JJ] Done.
hardware. This was aggravated by the increasing maintenance costs and extended downtime resulting from the continuous failures of existing clusters.
Considering this, we leveraged our experience in the design of web
applications, including JavaScript, since their inception and other
volunteer and 
unconventional distributed evolutionary computing systems, and we 
decided to design and release a new free framework that would allow anyone to
create a volunteer distributed EC experiment using cloud resources as
servers and browsers as clients.

@@

@@

Figure \ref{fig:system} describes the general system architecture and
algorithm behavior. Different web technologies, such as JQuery or chart.js have
been used to build the elements of the system. 
%FERGU: Mario, explica en el texto de abajo qué es winston en la figura también. 
%Vamos, que todo lo que salga en la figura tiene que ir en el texto en alguna parte :)
% [Mario] Ok voy a explicar la figura. Quería saber primero su opinión tal vez se les hacía mucho detalle.
% Creo que esta figura va mejor en la sección donde hablamos de
% webworkers. La anterior era más adecuada como visión general. [JJ]
\begin{figure}[!t]
\centering
\includegraphics[width=4in]{img/system.pdf}
\caption{Description of the proposed system. Clients execute a JS EA
  in the browser, which, every 100 generations, sends the best
  individual and receives a random one from the server. }  
\label{fig:system}
\end{figure}

@@

\begin{figure}[!t]
\centering
\includegraphics[width=12cm]{img/baseline-times.png}
\caption{Comparison of time to solution for the baseline system, using a {\tt
    node.js} client, for different population sizes. They only compare on
runs in which solution has been found.}
\label{fig:baseline}
\end{figure}
First we establish the baseline performance by running the
evolutionary algorithm in a desktop client written using NodEO, 
%FERGU: es NodEO o NodIO? Quito el \sf, por cierto, que queda feo
%Por qué usamos NodEO aquí y no NodIO, vamos a compararlos? 
%Qué diferencias hay entre los dos? (ponerlo en el SOA)
%[Mario] Creo que es NodEO y NodIO fué mi error.
the node.js evolutionary algorithm library. This experiment tries to 
%FERGU: quitar el "the node.js evolutionary algorithm library", si ya lo hemos dicho
%o explicarlo mejor: "the basic js library we have used as base to built NodIO", por ejemplo
find the solution to the 40-trap function with parameters $l=4, a=1,
b=2, z=3$, a population size equal 512, which was found in previous
experiments to be the correct one for finding the solution 95\% of the
%Fergu: cita a esos experimentos?
time. The algorithm was run until the solution (a string with all
ones) was found or five millions evaluations had been performed. It
took around a minute, on average, that is $68.9694$ seconds, to
perform the fifty runs. In this experiment only 33, that is, 66\% were
successful. The experiments were repeated for population $p=1024$,
with success rate upgraded to 100\% and average duration $3.46$
seconds. Results for these two experiments 
are shown in Figure \ref{fig:baseline}.

@@

Initial experiments were set up using the OpenShift
Platform-as-a-Service, which provides a free tier within which this
experiment could be performed. That way, and as required, the hardware
and cloud cost for this experiment were zero. Experiments were
announced through a post in Twitter and other social networks, and
results were published here \cite{DBLP:conf/gecco/GuervosG15}. For the
purpose of this paper, we repeated the announcement several times
through the months of April and then by the beginning of August. All
in all, we have the set of runs with the characteristics shown in
Table \ref{tab:runs}. In general, every experiment took several
days. No particular care was taken about the time of the announcement
or the particular wording. Every {\em experiment} consisted in running
until the solution of the 40-trap problem was found. When the correct
solution was sent to the server, the counter was updated and the pool
of solutions reset to the void set. No special care was taken to wait
until all clients had finished, thus it might happen that, in fact,
the islands running in the browser {\em spill} from one experiment to
the next. However, previous experiments have proved that the influence
of these islands in the next experiment is, in fact negligible.
%
\begin{table*}d
\caption{Characteristics of the volunteer computing initial runs. \label{tab:runs}}
\begin{center}
\begin{tabular}{l|rr}
\hline
Date & \# experiments & Different IPs \\
\hline 
April 4th 4/4 & 57 & 191 \\
April 24th 4/24 &  231 & 559 \\
July 31th 7/31 & 97 & 179 \\
\hline 
\end{tabular}
\end{center}
\end{table*}
%
\begin{table*}
\caption{Summary of time per run, number of IPS and number of {\tt
    PUT}s per IP in the initial runs. \label{tab:summary:os}}
\begin{center}
\begin{tabular}{l|ccccc}
\hline
Date & Median \#IPs & Median time (s) & Median \#{\tt PUT}s & $<$ 69s & $<$ 3.46s \\
\hline 
4/4 & 5 & 2040 & 18 & 14.29\% & 5.36\% \\
4/24 &  5 & 732 & 11 & 47.39\% & 3.91\% \\
7/31 & 5 & 260 & 23 & 27.08\% & 1.04\%  \\
\hline 
\end{tabular}
\end{center}
\end{table*}
%
The table shows that every run included more than 50 experiments. The
number of different IPs intervening in them varied from more than one
hundred to more than five hundred in the second experiment.

A summary of the results of each run is also shown in Table
\ref{tab:summary:os}. This table shows the median number of IPs
intervening in each experiment,  median of the time needed
to finish the experiment, median number of HTTP {\tt PUT}s per IP, and
then two columns showing the percentage of experiments that took less
than the two baseline experiments shown in the introduction to this
section. The first striking result is that in all cases, 50\% of the
experiments involved 5 or less IPs. This is consistent with results
found previously \cite{DBLP:conf/gecco/GuervosG15} which found 6 to be
the usual number of IPs that participated in an experiment.

\subsection{{\sf NodIO-W$^2$}, a WebWorker-based architecture}

\begin{figure}[!t]
\centering
\includegraphics[width=4in]{img/Algorithm.pdf}
\caption{Description of the W$^2$ version of {\sf NodIO}. In this
  version, clientts use WebWorkers to run the evolutionary algorithm,
  with two of them per browser.}  
\label{fig:system:w2}
\end{figure}
